\section{Discussion}

\subsection{AI Framework Consistency}
Our framework demonstrates robust performance in automating the construction of complex ecosystem models across diverse marine environments. The successful processing of over 41,000 species across three distinct regions validates the framework's scalability and broad applicability. The framework's ability to maintain consistent species classifications while adapting to regional ecological differences suggests it effectively captures fundamental ecological relationships.

The computational efficiency analysis reveals important insights about framework scalability. Data harvesting and diet collection emerge as the primary computational bottlenecks, particularly evident in the South East Inshore region's extended processing times. These bottlenecks likely stem from API rate limitations and the complexity of extracting ecological information from diverse data sources. The relatively constant processing times for species identification, grouping, and parameter estimation across regions indicate these components scale efficiently with increasing species counts.

The framework's classification consistency merits particular attention. The high stability scores (98.8-99.6\%) across regions demonstrate reliable species-to-group assignments. The systematic nature of classification inconsistencies provides valuable ecological insights. Species with ambiguous classifications often represent organisms that naturally span multiple ecological niches. For instance, the alternating classifications of anemones between benthic infaunal carnivores and filter feeders reflect their complex feeding strategies. Similarly, the variable classification of flatfishes between benthivore and demersal categories aligns with their known ecological plasticity. These classification patterns suggest the framework captures meaningful ecological uncertainty rather than arbitrary assignment errors.

The diet matrix validation reveals a nuanced picture of trophic relationship stability. The moderate negative correlations between iterations initially appear concerning. However, the high stability of significant feeding interactions (89.8-92.5\%) suggests the framework maintains consistent broad-scale trophic structure while allowing flexibility in fine-scale interactions. This pattern aligns with ecological theory, where core trophic relationships remain stable while peripheral feeding interactions may vary with resource availability and environmental conditions.

The observed trophic level patterns provide compelling evidence for the framework's ecological validity. The consistent classification of higher trophic level species reflects the relatively constrained niches of specialized predators. Conversely, the greater variability in lower trophic level classifications mirrors the natural complexity and adaptability of these groups. The framework's ability to capture this fundamental ecological pattern suggests it successfully incorporates biological realism into its classification decisions.

The regional differences in group size variation and classification stability offer insights into ecosystem complexity. The Northern Territory's higher variation in group sizes and slightly lower classification stability likely reflect the increased ecological complexity of tropical reef systems. The more stable classifications in the South East regions may indicate more clearly defined ecological niches in temperate marine environments. These regional patterns demonstrate the framework's sensitivity to underlying ecological differences while maintaining consistent overall performance.

The reduction from 63 potential functional groups to 34-36 region-specific groups indicates the framework's ability to identify ecologically relevant groupings while avoiding artificial complexity. The statistical consistency across regions suggests these groupings represent fundamental ecological units rather than arbitrary divisions. This optimization of functional group complexity balances model detail with practical utility, a crucial consideration for ecosystem modeling applications.


\subsection{Limitations and Uncertainties}

Our framework faces several AI-specific limitations identified in recent ecological modeling research. The framework's performance depends on the quality and completeness of available training data. The framework's classification patterns showed regional variations in stability, though further research is needed to determine the relationship between data availability and classification performance. This consideration aligns with findings from \cite{Kuhn2024} regarding machine learning applications in fisheries, where data quality significantly impacts model reliability.

A significant limitation of our approach stems from its reliance on Claude 3.5 Sonnet, a closed-source large language model. The proprietary nature of this model introduces uncertainty regarding the training data used in its development and potential biases that may affect ecological interpretations. While our validation demonstrates consistent performance, the inability to examine the model's training data or internal decision-making processes raises important considerations for scientific reproducibility. Future iterations of the framework may benefit from exploring open-source alternatives or implementing multiple model approaches as demonstrated by \cite{Kommineni2024} in their work with various LLMs for biodiversity research.

The framework's interpretability presents another key challenge. While our validation demonstrates robust performance metrics, the underlying AI decision-making processes, particularly in parameter estimation, require careful scrutiny. This challenge mirrors concerns raised by \cite{Fernandes2024} regarding the "black box" nature of AI systems in aquaculture applications. Our framework partially addresses these concerns through explicit uncertainty quantification and validation metrics, but further work is needed to enhance model transparency.

Technical limitations include computational resource requirements and processing time constraints, particularly evident in data harvesting operations. These limitations align with implementation barriers identified by \cite{Fernandes2024}, including acquisition costs and technical expertise requirements. The framework's sensitivity to data availability varies across ecological roles and regions, affecting both classification stability and diet matrix reliability.

\subsection{Applications for EBFM}

The framework offers significant practical value for ecosystem-based fisheries management through several key capabilities. Managers can now construct initial EwE models for new regions in days rather than months, enabling faster response to management needs. This addresses a key bottleneck identified by \cite{Zheng2023} in marine resource management. The framework's explicit quantification of uncertainty in species classifications and trophic relationships enables managers to identify areas requiring additional data collection or careful monitoring, following approaches recommended by \cite{Kuhn2024}. 

The demonstrated adaptability across diverse ecosystems enables customization for specific regions while maintaining methodological consistency, supporting standardized approaches to ecosystem management across jurisdictions. Through systematic processing of available data, the framework also reveals specific areas where additional research or monitoring would most improve model reliability, aligning with recent work by \cite{Chen2024} on marine protected area management.

\subsection{Future Directions}

Future development should focus on several key areas to enhance the framework's utility and reliability. Integration with emerging marine AI systems, such as specialized models like MarineGPT \citep{Zheng2023} and multimodal frameworks like LITE \citep{Li2024}, could enhance the framework's capabilities in processing region-specific information and handling incomplete data. The LITE framework's success in combining semantic time-series with temporal trend images suggests promising directions for improving our framework's robustness to data gaps and distribution shifts. Their sparse Mixture-of-Experts approach to handling incomplete features could particularly benefit our diet matrix construction process. Implementation of standardized evaluation frameworks, such as those proposed in the ELLE dataset \citep{Guo2025}, would enable more rigorous assessment of the framework's performance across different ecological contexts. Development of improved visualization and explanation tools for AI decision-making processes would address current limitations in model transparency and support broader adoption in management contexts.

\subsection{Conclusion}

Our validation analysis demonstrates both the capabilities and limitations of AI-assisted ecosystem modeling. The framework shows remarkable stability in many aspects while highlighting areas of ecological uncertainty that deserve attention. The clear regional patterns in performance suggest that the approach can adapt to different ecological contexts while maintaining scientific rigor. These findings support the framework's utility for ecosystem-based management while providing clear directions for future improvement.

The observed trade-offs between consistency and complexity reflect fundamental challenges in ecosystem modeling rather than simple methodological limitations. By quantifying these trade-offs and their regional variations, our analysis provides a foundation for more informed application of ecosystem models in fisheries management. As marine ecosystems face increasing pressures from climate change and human activities, this understanding of model behavior and limitations becomes increasingly crucial for effective ecosystem-based management \citep{Geary2020}.
