\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage[left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage[round,authoryear]{natbib}  % Specify author-year citation style
\usepackage{hyperref}
\usepackage{xr}                % For cross-referencing between documents
\usepackage{cleveref}          % For intelligent cross-referencing
\usepackage{pdflscape}         % For landscape pages
\usepackage{mdframed}          % For creating boxed environments
\usepackage{xcolor}            % For custom colors
\usepackage{tikz}             % For creating diagrams and flowcharts
\usetikzlibrary{positioning}  % For relative positioning in TikZ
\usetikzlibrary{shapes.geometric}  % For geometric shapes in TikZ
\usepackage{lineno}

% Define custom prompt environment
\definecolor{promptbg}{RGB}{250,250,250}
\definecolor{promptframe}{RGB}{200,200,200}
\newmdenv[
    linewidth=1pt,
    backgroundcolor=promptbg,
    linecolor=promptframe,
    leftmargin=1em,
    rightmargin=1em,
    innerleftmargin=1em,
    innerrightmargin=1em,
    innertopmargin=1em,
    innerbottommargin=1em,
    skipabove=1em,
    skipbelow=1em
]{prompt}

% Configure hyperref after other packages
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,
    urlcolor=cyan,
    citecolor=blue
}

\title{Automated Diet Matrix Construction for Marine Ecosystem Models Using Generative AI}

\author{Scott Spillias\textsuperscript{1,2}\thanks{Corresponding author: scott.spillias@csiro.au} \and
Beth Fulton\textsuperscript{1,2} \and
Fabio Boschetti\textsuperscript{2,3} \and
Joanna Strzelecki\textsuperscript{3} \and
Rowan Trebilco\textsuperscript{1,2}}

\date{\today}  % Use current date or remove for no date

% Author affiliations
\newcommand{\affiliations}{
\noindent\textsuperscript{1}CSIRO Environment, Hobart, Australia\\
\textsuperscript{2}Centre for Marine Socio-Ecology, University of Tasmania, Hobart, Australia\\
\textsuperscript{3}CSIRO Environment, IOMRC Crawley, Australia\\
\textsuperscript{4}CSIRO Environment, St. Lucia, Australia
}

\begin{document}
\linenumbers
\maketitle
\affiliations

\begin{abstract}
This study introduces and validates a novel AI-driven framework for automated species grouping and diet matrix generation in Ecopath with Ecosim (EwE) ecosystem models, addressing a critical bottleneck in model development. We evaluate the framework across three contrasting Australian marine regions (Shelf and Offshore regions of South-East Tasmania and the Northern Territory), processing over 41,000 species through multiple validation iterations. The framework successfully condensed 63 potential functional groups into 34-36 region-specific groups, achieving high classification stability (>99\% consistency) for most species. Notably, the framework demonstrated robust performance in the South East offshore region with only 0.03\% inconsistent classifications, while showing greater variability in complex tropical systems (1\% inconsistent classifications). Higher trophic level species maintained consistent classifications across all runs, with the framework identifying 235-327 significant predator-prey interactions per region at >70\% consistency. This systematic validation reveals that while the framework can reliably automate species grouping, its performance varies predictably with ecosystem complexity and data availability. These findings provide quantitative evidence for the framework's ability to create repeatable components for ecosystem model development. Our results demonstrate the potential for AI to significantly reduce model development time, offering a practical pathway to expand the application of ecosystem-based management across diverse marine environments. We offer a pathway towards further validation of this framework to ensure ecological accuracy.
\end{abstract}

\input{introduction}
\input{methods}
\input{results}
\input{discussion}

\section*{Acknowledgements}
[Add acknowledgements here]

\section*{Data Availability}
The complete codebase, including all scripts, configuration files, and analysis tools, is available at [GitHub repository URL]. The validation framework, including reference group definitions and classification rules, is documented in the project repository to ensure reproducibility.

\section*{Author Contributions}
SS: Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Resources, Data Curation, Writing - Original Draft, Writing - Review \& Editing, Visualization, Project administration. BF: Validation, Writing - Review \& Editing, Supervision, Funding acquisition. FB: Methodology, Software, Validation, Writing - Review \& Editing, Supervision. CB: Investigation, Validation, Writing - Review \& Editing. RB: Investigation, Validation, Writing - Review \& Editing. JPG: Investigation, Validation, Writing - Review \& Editing. JS: Conceptualization, Validation, Investigation, Writing - Review \& Editing. RS: Investigation, Validation, Writing - Review \& Editing. RT: Methodology, Software, Validation, Investigation, Writing - Review \& Editing, Supervision, Funding acquisition.

\section*{Statement on the Use of Generative AI}
Generative AI tools, specifically Claude Sonnet 3.5, were utilized in the preparation of this manuscript to assist with tasks such as language refinement, text structuring, and summarization. All scientific content, data interpretation, and conclusions were independently developed and verified by the authors to ensure accuracy and integrity.

\bibliographystyle{apalike}
\bibliography{references}

\clearpage
\input{supplement}

\end{document}
