\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{palatino}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage[left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[RO,LE]{Diet Matrices with Generative AI}

\title{Automated Diet Matrix Construction for Marine Ecosystem Models Using Generative AI}

\author{Scott Spillias\textsuperscript{1,2}\thanks{Corresponding author: scott.spillias@csiro.au} \and
Elizabeth A. Fulton\textsuperscript{1,2} \and
Fabio Boschetti\textsuperscript{2,3,4} \and
Cathy Bulman\textsuperscript{1} \and
Joanna Strzelecki\textsuperscript{3} \and
Rowan Trebilco\textsuperscript{1,2}}

\date{\today}

% Author affiliations
\newcommand{\affiliations}{
\noindent\textsuperscript{1}CSIRO Environment, Hobart, Australia\\
\textsuperscript{2}Centre for Marine Socio-Ecology, University of Tasmania, Hobart, Australia\\
\textsuperscript{3}CSIRO Environment, IOMRC Crawley, Australia\\
\textsuperscript{4}Oceans Institute, The University of Western Australia, Crawley, WA\\
}

\begin{document}
\maketitle
\affiliations

\section*{Acknowledgements}
SS was funded by a CSIRO R+ Postdoctoral Fellowship. This work was partially supported by the 'Futures of Seafood' project.

\section*{Data Availability}
The complete codebase, including all scripts, configuration files, and analysis tools, is available at https://github.com/s-spillias/AI-EwE-Diets. The validation framework, including reference group definitions and classification rules, is documented in the project repository to ensure reproducibility.

\section*{Author Contributions}
SS: Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Resources, Data Curation, Writing - Original Draft, Writing - Review \& Editing, Visualization, Project administration. BF: Validation, Writing - Review \& Editing, Supervision, Funding acquisition. FB: Methodology, Software, Validation, Writing - Review \& Editing, Supervision. CB: Investigation, Data Curation, Validation, Writing - Review \& Editing. JS: Conceptualization, Validation, Investigation, Writing - Review \& Editing. RT: Methodology, Software, Validation, Investigation, Writing - Review \& Editing, Supervision, Funding acquisition.

\section*{Statement on the Use of Generative AI}
Generative AI tools, specifically Claude Sonnet 3.5, were utilized in the preparation of this manuscript to assist with tasks such as language refinement, text structuring, and summarization. All scientific content, data interpretation, and conclusions were independently developed and verified by the authors to ensure accuracy and integrity.

\end{document}